Large squeezing in multicomponent trapped Bose-Einstein condensates

### 2012-11-27 large_number_squeezing_with_atoms.tex
Add packages so it actually compiles

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass{iopart}
%\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}
\usepackage{graphicx}
\usepackage{ulem}
\usepackage{amssymb}

@@

We integrate equation (\ref{eqStochasticEquations}) numerically in one, two and three spatial dimensions using the numerical package XMDS \cite{Dennis2012}.  The dimensional reduction is achieved by \textbf{Derive link between mode shape, nonlinearity U, and the SM equivalent nonlinearity, so we can always compare a given high-dimensional multimode sim against the equivalent ideal analytic single mode case}

\section{Effects of trap geometries on squeezing in 1D}
\begin{itemize}
  \item Plot of squeezing for given set of params for gaussian, TF, constant density, show it get steadily better
\end{itemize}

Some common mode shapes are the groundstates of a harmonic trap with negligible nonlinearity (a gaussian), a harmonic trap with strong nonlinearity (Thomas-Fermi), and a constant potential (constant density). Using (\ref{eqChiUequivalence}) we find the following equivalences
\begin{eqnarray}
\chi &=& U \sqrt{\frac{m^3 \omega_x \omega_y \omega_z}{8 \pi^3 \hbar^3}} \,\,\,\, ({\textrm{Gaussian}}) \\
%
\chi &=& \frac{4}{7} \left( \frac{15 U \omega_x \omega_y \omega_z}{16 \pi \sqrt{2}} \right)^{2/5} \left( \frac{m}{N} \right)^{3/5} \,\,\,\, ({\textrm{Thomas-Fermi}}) \\
%
\chi &=& \frac{U}{V} \,\,\,\, {\textrm{(constant density)}}
\end{eqnarray}
Choosing parameters where the mode shape of the multimode system doesn't change much during the evolution allows us to compare the effect of mode shape on the best squeezing that can be attained relative to the two mode system. An example in shown in Figure~\ref{figModalSqueezingEffects1D}. Clearly the closer the mode shape is to constant density, the better.

\begin{figure}
    \centering
    \includegraphics[width=8cm]{figures/modal_squeezing_effects_1D.eps}
    \caption{Effects of mode shape on squeezing. The plot shows the normalized number variance in mode $\hat{a}$ as a function of the time the final coupling pulse is applied. Thick red line, analytic two-mode solution; solid blue line, constant density mode; dashed line, Thomas-Fermi mode; dash-dotted line, gaussian mode. The nonlinearities of the various 1D simulations have been adjusted for their mode shape so that they are equivalent to that of the 0D, two-mode mode nonlinearity. Parameters: $n_a = n_b =1 \times 10^5$, $\chi_{aa}=0.03 \hbar, \chi_{ab}=\chi_{bb}=0, \tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=4.0$. }
    \label{figModalSqueezingEffects1D}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing.eps}
    \label{figDimensionalSqueezingEffects:sub1}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing.eps}
    \label{figDimensionalSqueezingEffects:sub2}
  \end{subfigure}
\caption{Effects of dimension on squeezing in two different nonlinear regimes. If two nonlinearities are non-zero, the deleterious effects of higher dimensions on squeezing are exacerbated. Thick red line, analytic two-mode solution; solid blue line, 1D; dashed line, 2D; dash-dotted line, 3D. (a) Parameters: $n_a = n_b =2 \times 10^5$, $\chi_{aa}=\chi_{bb}=0.03\hbar$, $\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.1$. (b) $n_a = n_b =2 \times 10^5$, $\chi_{aa}=0.03\hbar$, $\chi_{bb}=\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.1$.}
  \label{figDimensionalSqueezingEffects}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass{iopart}
%\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}
\usepackage{graphicx}
\usepackage{ulem}
\usepackage{amssymb}
\usepackage{caption}
\usepackage{subcaption}

@@

We integrate equation (\ref{eqStochasticEquations}) numerically in one, two and three spatial dimensions using the numerical package XMDS \cite{Dennis2012}.  The dimensional reduction is achieved by estimating the mode shape in three dimensions, and using dimensionally reduced values for $U_{ij}$ that match the zero-dimensional reduction given by Eq.~(\ref{eqChiUequivalence}).  An equivalent process is to match the chemical potentials of the initial states.

\section{Effects of trap geometries on squeezing in 1D}

A one-dimensional simulation is sufficient to examine the hypothesis that the squeezing will be higher for a BEC where the spatial mode has a more constant density profile.  We compare squeezing for fields starting in the ground states of a harmonic trap with negligible nonlinearity (a Gaussian), a harmonic trap with strong nonlinearity (Thomas-Fermi), and a constant potential (constant density). Using (\ref{eqChiUequivalence}) we find the following equivalences
\begin{eqnarray}
\chi &=& U \sqrt{\frac{m^3 \omega_x \omega_y \omega_z}{8 \pi^3 \hbar^3}} \,\,\,\, ({\textrm{Gaussian}}) \\
%
\chi &=& \frac{4}{7} \left( \frac{15 U \omega_x \omega_y \omega_z}{16 \pi \sqrt{2}} \right)^{2/5} \left( \frac{m}{N} \right)^{3/5} \,\,\,\, ({\textrm{Thomas-Fermi}}) \\
%
\chi &=& \frac{U}{V} \,\,\,\, {\textrm{(constant density)}}
\end{eqnarray}
\textbf{Choosing parameters where the mode shape of the multimode system doesn't change much during the evolution allows us to compare the effect of mode shape on the best squeezing that can be attained relative to the two mode system.  Joe: We need to be more specific here.}

The results of this simulation are shown in Figure~\ref{figModalSqueezingEffects1D}, which clearly confirms the hypothesis that . Clearly the closer the mode shape is to constant density, the better.

\begin{figure}
    \centering
    \includegraphics[width=8cm]{figures/modal_squeezing_effects_1D.eps}
    \caption{Effects of mode shape on squeezing. The plot shows the normalized number variance in mode $\hat{a}$ as a function of the time the final coupling pulse is applied. Thick red line, analytic two-mode solution; solid blue line, constant density mode; dashed line, Thomas-Fermi mode; dash-dotted line, Gaussian mode. The nonlinearities of the various 1D simulations have been adjusted for their mode shape so that they are equivalent to that of the 0D, two-mode mode nonlinearity. Parameters: $n_a = n_b =1 \times 10^5$, $\chi_{aa}=0.03 \hbar, \chi_{ab}=\chi_{bb}=0, \tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=4.0$. }
    \label{figModalSqueezingEffects1D}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing_1.eps}
    \label{figDimensionalSqueezingEffects:sub1}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=7cm]{figures/dimensional_effects_on_squeezing_2.eps}
    \label{figDimensionalSqueezingEffects:sub2}
  \end{subfigure}
\caption{Effects of dimension on squeezing in two different nonlinear regimes. If two nonlinearities are non-zero, the deleterious effects of higher dimensions on squeezing are exacerbated. Thick red line, analytic two-mode solution; solid blue line, 1D; dashed line, 2D; dash-dotted line, 3D. (a) Parameters: $n_a = n_b =2 \times 10^5$, $\chi_{aa}=\chi_{bb}=0.03\hbar$, $\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.1$. (b) $n_a = n_b =2 \times 10^5$, $\chi_{aa}=0.03\hbar$, $\chi_{bb}=\chi_{ab}=0$, $\tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=3.1$.}
  \label{figDimensionalSqueezingEffects}
\end{figure}
